### 2018-07-16 manuscript.rmd
incorporating data and initial data wrangling chunks

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

>>>>

```{r setup, include=FALSE}
	# Load packages
	## General
library(knitr)
library(tidyverse)
library(reshape2)
	## Graphics
library(FactoMineR)
library(factoextra)
	## Biological
library(vegan)
library(ape)
library(phangorn)
library(phytools)
library(OUwie)
library(picante)
library(geiger)
library(phylobase)
library(fields)
library(phylosignal)
library(geomorph)
library(phylopath)
library(Rphylip)
library(arbutus)

	# Configure knitr, see http://yihui.name/knitr/options
	opts_knit$set(
	  progress=TRUE,
	  verbose=TRUE)
	opts_chunk$set(
	#  include=FALSE,
	  cache=TRUE,
	  echo=FALSE,
	  message=FALSE
	 )
	# Set paths to input data
	setwd("~/tentilla_morph")

	#Load raw data
  read.csv("byslide.csv") -> numbers
  categorical <- read.csv("Homolog_Categorical.csv")[,-2]
  rownames(categorical) = categorical$Species

  #Polish fields
  castnumbers = numbers
  castnumbers[castnumbers=="needDIC"] <- NA
  castnumbers[castnumbers=="needConfocal"] <- NA
  castnumbers[castnumbers=="needMature"] <- NA
  castnumbers[castnumbers=="needReslide"] <- NA
  castnumbers[castnumbers==-1] <- NA
  castnumbers[,c(-1,-2,-3)] <- sapply(castnumbers[,c(-1,-2,-3)],as.character)
  castnumbers[,c(-1,-2,-3)] <- sapply(castnumbers[,c(-1,-2,-3)],as.numeric)

  #Define functions to deal with NAs and decimals
  mean.na <- function(a){mean(a, na.rm = TRUE)}
  var.na <- function(a){var(a, na.rm = TRUE)}
  round3 <- function(a){round(a, 3)}

  #Compute morphometric ratios
  coiledness = castnumbers$Cnidoband.free.length..um./castnumbers$Cnidoband.length..um.
  heteroneme_elongation = castnumbers$Heteroneme.free.length..um./castnumbers$Heteroneme.width..um.
  haploneme_elongation = castnumbers$Haploneme.free.length..um./castnumbers$Haploneme.width..um.
  desmoneme_elongation = castnumbers$Desmoneme.length..um./castnumbers$Desmoneme.width..um.
  rhopaloneme_elongation = castnumbers$Rhopaloneme.length..um./castnumbers$Rhopaloneme.width..um.
  heteroneme_shaft_extension = castnumbers$Heteroneme.free.length..um./castnumbers$Heteroneme.shaft.free.length..um.
  Heteroneme_to_CB = castnumbers$Heteroneme.free.length..um./(castnumbers$Cnidoband.free.length..um.+0.001)
  total_heteroneme_volume = castnumbers$Heteroneme.volume..um3.*castnumbers$Heteroneme.number
  total_haploneme_volume = (castnumbers$Cnidoband.free.length..um./castnumbers$Haploneme.width..um.)*((4*pi/3)*0.5*castnumbers$Haploneme.free.length..um.*((0.5*castnumbers$Haploneme.width..um.)^2))
  cnidomic_index = log(total_heteroneme_volume + total_haploneme_volume)
  cnidomic_index[which(is.na(cnidomic_index))] <- log(total_haploneme_volume[which(is.na(cnidomic_index))])
  cnidomic_index[which(is.na(cnidomic_index))] <- log(castnumbers[which(is.na(cnidomic_index)), "Heteroneme.volume..um3."])
  cnidomic_index[cnidomic_index==-Inf]<-NA
    morphometrics = data.frame(castnumbers$slide_id, castnumbers$Species,coiledness, heteroneme_elongation, haploneme_elongation, desmoneme_elongation, rhopaloneme_elongation,heteroneme_shaft_extension, Heteroneme_to_CB, total_heteroneme_volume, total_haploneme_volume, cnidomic_index)
  names(morphometrics)[1:2] = c("slide_id"," Species")
  castnumbers = data.frame(castnumbers,morphometrics[,c(-1,-2)])

  #logtransform variables which are not normal
  castlogs = castnumbers
  non_normal = apply(castnumbers[,c(-1,-2,-3)], 2, shapiro.test) %>% lapply(function(x){x$p.value}) %>% unlist() 
  non_normal = which(as.vector(non_normal) < 0.05)+3
  castlogs[,non_normal] <- sapply(castnumbers[,non_normal], log)
  castlogs[castlogs==-Inf]<-NA

  #Means, varinces, standard errors
  castmeans <- aggregate(. ~  Species, data = castnumbers[,c(-1,-3)], mean.na, na.action = na.pass)
  castmean_logs <- aggregate(. ~  Species, data = castlogs[,c(-1,-3)], mean.na, na.action = na.pass)
  castvariances <- aggregate(. ~  Species, data = castnumbers[,c(-1,-3)], FUN = var.na, na.action = na.pass) #[,-2]
  castvariance_logs <- aggregate(. ~  Species, data = castlogs[,c(-1,-3)], FUN = var.na, na.action = na.pass)
  castvariances[is.na(castvariances)] <- 0
  castvariance_logs[is.na(castvariance_logs)] <- 0
  castN <- aggregate(. ~  Species, data = castnumbers[,c(-1,-3)], length, na.action = na.pass) 
  castN = data.frame(castN$Species, rowMeans(castN[,-1])) #[,-2]
  names(castN) = c("Species", "N_specimens")
  castSE <- cbind(castvariances$Species, sqrt(castvariances[,-1])/sqrt(castN$N_specimens))
  castSE_logs <- cbind(castvariance_logs$Species, sqrt(castvariance_logs[,-1])/sqrt(castN[,-1]))
  names(castSE)[1] <- "Species"
  names(castSE_logs)[1] <- "Species"
```
